

## Submit

I care more about you attempting these than 100% correct answers. 

To submit, generate your answer below, and paste it into the google form provided. 

If you're in DSI 36100 (which you probably are), that's https://forms.gle/rYs9hhL726zmE9wNA 

```{r passr, message=FALSE, warning=FALSE, include=FALSE}
#this file gets included/sourced/child rmd into each learnr Rmd. It means I can update how things get submitted on all tutorials without needing to manually edit them all.

#see passr project file for ideas for other ways to do this (email, gdrive distrib, etc.)

library(learnrhash)
```


```{r encode-me, context="server"}

learnrhash::encoder_logic()

#learnSTATS::encoder_logic()
```

### Hash

Use the copy and paste function to submit this to the form

```{r encode-ui, echo=FALSE}

learnrhash::encoder_ui(ui_before = NULL, ui_after = NULL)

```

### Decode

If you want to check what the hash shows, use the decoder (at the moment this is outputting in the instructions.Rmd chunk output where you can easily view it; If you run the tmpReport.Rmd it shows up within the rendered shiny...I'm not sure why)

```{r context="server"}
learnrhash::decoder_logic()
```


```{r decode-ui, echo=FALSE}
learnrhash::decoder_ui()
```


